---
title: "Homework 3"
author: "Jerry Slutsky"
date: "2/16/2020"
output: 
  html_document:
    keep_md: true
    code folding: hide
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error = FALSE)
library(psych)
library(tidyverse)
library(cowplot)
library(ggplot2)
library(ggthemes)
library(ggcorrplot)
```


# Overview

*This code is used for wrangling and visualizing the psych::bfi data set*
  
  
**First I will create an 'id' column for each individual in the dataset:**
```{r}
data<- bfi
data<- rownames_to_column(data, var = "id")
names(data)
```

  
## Task Option 8:

### Create composite scores for the Big Five scores

```{r}
new.data<- data %>% 
  as_tibble() %>% 
  mutate(A_sum = rowSums(select(., A1:A5),na.rm = TRUE)) %>% 
                mutate(N_sum = rowSums(select(., N1:N5),na.rm = TRUE)) %>% 
                mutate(O_sum = rowSums(select(., O1:O5),na.rm = TRUE)) %>% 
                mutate(C_sum = rowSums(select(., C1:C5),na.rm = TRUE)) %>% 
                mutate(E_sum = rowSums(select(., E1:E5),na.rm = TRUE))
```


**Describe relationships among composite (sum scores) of the Big 5 Personality traits**

  * I used the 'psych' package to create a correlation heatmap. 

```{r}
corcomp <- corFiml(new.data[c(30:34)])
cor.plot(corcomp, upper = FALSE, numbers = TRUE, show.legend = TRUE)
```

The results suggest that the Big 5 composite have small relationships overall. The largest correlation is between openness and concientiousness (r = .28) and the smallest correlation is between neuroticism and extraversion (r = .05)


**Graph of Big 5 Composite Scores**

  * Created a barplot of all the Big 5 means
```{r}
fun_mean <- function(x){
  return(data.frame(y=round(mean(x),2),label=round(mean(x,na.rm=T),2)))}
new.data %>%
  gather(A_sum, N_sum, O_sum, C_sum, E_sum, key = "Trait", value = "Composite") %>%
  ggplot(aes(x=Trait, y=Composite, fill=Trait)) +
  geom_bar(stat="summary", width = 0.5, fun.y = "mean")+
  stat_summary(fun.y = mean, geom="point",colour="black", size=1) +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7) +
  labs(x="Big 5 Traits",y="Composite scores")
```

Agreeableness has the highest composite mean and neuroticism has the composite lowest mean


## Task Option 9:

### Does personality appear to change for different ages?


**Used facet_wrap to graph all 5 composites in one figure**

```{r, fig.align='center'}
new.data %>%
  gather(A_sum, N_sum, O_sum, C_sum, E_sum, key = "Trait", value = "Composite") %>%
  ggplot(aes(x = age, y = Composite, color = Trait)) +
    geom_smooth(method="lm", col="black", fun.y = "mean") +
    geom_point(alpha= .5) +
    facet_wrap(~ Trait, scales = "free") +
    theme_bw() + theme(legend.position="none") +
   labs(x="Age", y="Composite Score")
```

Looks like neuroticism and agreeableness are the only personality traits that change over time according to these data



## Task Option 10

### Explore how gender is related to the different personality variables.


**Making male = 1 and female = 2

```{r}
new.data$gender<- factor(new.data$gender, levels=c(1,2), labels=c("Male","Female"))
```


**Using the describeBy function to get descriptives for each trait**
```{r}
A<-describeBy(new.data$A_sum, group = new.data$gender)
C<-describeBy(new.data$C_sum, group = new.data$gender)
N<-describeBy(new.data$N_sum, group = new.data$gender)
O<-describeBy(new.data$O_sum, group = new.data$gender)
E<-describeBy(new.data$E_sum, group = new.data$gender)
```

**Pull means out of the describeBy function from above:**
```{r, error=FALSE, warning=FALSE}
males<- c(N$`Male`$mean, A$`Male`$mean, C$`Male`$mean, O$`Male`$mean, E$`Male`$mean) 
females<- c(N$`Female`$mean, A$`Female`$mean, C$`Female`$mean, O$`Female`$mean, E$`Female`$mean) 
```

**Combine Means and Adding Row Names**
```{r}
means<-data.frame(males,females, check.names = FALSE, 
                  check.rows = FALSE,
                  row.names=c("N","A","C","O","E"))
```

**Printing Means of Males and Females in Table Format:**
```{r, fig.align='center'}
knitr::kable(means)
```

As you can see in the table above containing means between gender for each of the Big 5 traits from the BFI data set, Males scored higher on Conscientiousness, Openness, and Extroversion. Females scored higher on Neuroticism and Agreeableness.
