---
title: "hw2"
author: "Jerry"
date: "February 8, 2020"
output: 
  html_document:
    keep_md: true
    theme: paper
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, warning = FALSE, message = FALSE}
library(gapminder)
library(tidyverse)
```



Exercise 1

1.1
Use filter() to subset the gapminder data to three countries of your choice in the 1970's.

```{r}
gapminder %>%
  filter(year < 1980 & year > 1970, country %in% c("Afghanistan", "Albania", "Algeria"))
```


1.2
Use the pipe operator %>% to select "country" and "gdpPercap" from your filtered dataset in 1.1.

```{r}
gapminder %>%
  filter(year < 1980 & year > 1970, country %in% c("Afghanistan", "Albania", "Algeria")) %>% 
  select('country', 'gdpPercap')
```

1.3
Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that's the increase in life expectancy in your tibble.

Hint: you might find the lag() or diff() functions useful.

```{r}
library(tidyr)
```


```{r}
life_dec <- gapminder %>% 
  arrange(year) %>% 
  group_by(country) %>%
  mutate(lifeExp_dec = lifeExp - lag(lifeExp))
print(life_dec)

life_dec %>% 
  tidyr::drop_na()
```




1.4
Choose one of the following:

Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.


Hint: you might find the sort() function useful, or perhaps even the dplyr::slice() function.
  
  
```{r}
gapminder %>%
group_by(country) %>%
summarise(max(gdpPercap))
```



1.5
Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. Ensure GDP per capita is on a log scale.

```{r}
gapminder %>%
  filter(country %in% c("Canada")) %>% 
  select('lifeExp', 'gdpPercap') %>% 
  ggplot(gapminder, mapping = aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.5) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  theme_bw() +
  ylab("Life Expectancy")
```


Exercise 2
Pick one categorical variable and one quantitative variable to explore. Answer the following questions in whichever way you think is appropriate, using dplyr:

Range of values for life exepctancy.

```{r}
  range(gapminder$lifeExp)
```
List of call continents
```{r}
unique(gapminder$continent)
```


Mean life expectancy

```{r}
dplyr::summarise(gapminder, avg = mean(lifeExp))
```
Standard deviation of life expectancy

```{r}
dplyr::summarise(gapminder, sd = sd(lifeExp))
```

Frequency of each continent
```{r}
dplyr::count(gapminder, continent)
```



Exercise 3
Make two plots that have some value to them. That is, plots that someone might actually consider making for an analysis. Just don’t make the same plots we made in class—feel free to use a data set from the datasets R package or any other source if you wish.

A scatterplot of two quantitative variables.
One other plot besides a scatterplot.
You don’t have to use all the data in every plot! It’s fine to filter down to one group or country or a small handful of groups/countries.

```{r}
library(datasets)
```

Frequency of temperature found in each month in NYC in the summer of 1973. 5 = May, 9 = September. Temperature is in feirenheit.

```{r}
  ggplot(airquality, mapping = aes(Month, Temp)) +
  geom_point(alpha = 0.5) +
  xlab("Month") +
  theme_bw() +
  ylab("Temperature")
```

Wind in mph taken from LaGuardia airport in 1973. Looks like June, July, and August are less windy months than May and September.

```{r}
airquality %>% 
  ggplot(aes(Month, Wind)) +
  aes(group = Month)+
  geom_boxplot()
```



Recycling (Optional) (extra 2%)
Rubrics:

Reasoning
How well did you reason through the scenario?
For people who want to take things further.

Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?

```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

No, it looks like it pulls data where Afghanistan and Rwanda are on unique rows. When there is crossover data (when they share the same year), data is excluded. The best way to fix this would probably be this syntax:

```{r}
gapminder %>%
    filter( country %in% c("Afghanistan", "Rwanda"))
```

